Extended Data

### 2022-01-24 doc/manuscript/Dow_manuscript_Nature.Rmd
added Tumajer et al. 2022 ref
@camerondow35 , FYI

While responses of leaf phenology and seasonal _NEE_ to warming spring temperatures have been documented[@jeong_phenology_2011; @keenan_net_2014; @buermann_widespread_2018; @ibanez_forecasting_2010; @menzel_european_2006; @menzel_growing_1999;@finzi_carbon_2020], little is known about how the longest-lived component of fixed C in trees, the woody growth, is responding to warming spring temperatures. 
In fact, we are aware of only two studies that have documented stem-growth phenology of temperate deciduous forests over multiple years[@dorangeville_peak_2021; @etzold_number_2022]. <!-- I'm not aware of any others, but we do want to make sure we're not missing anything. klesse_integrating_2016 and martin_hydrometeorology_2017 both do this, but for conifers.-->
**The climate sensitivity of woody growth phenology in temperate deciduous trees and its link to annual growth has never been studied *in-situ* (but see Ref.[@zohner_how_2021] for a controlled sapling experiment).** <!-- revise this statement. Etzold et al kind of do this.-->

@@

Using dendrometer band measurements taken throughout multiple growing seasons at the Smithsonian Conservation Biology Institute (SCBI; Virginia, USA; n = `r n_dendro_trees_SCBI` trees from 2011-2020) and Harvard Forest (Massachusetts, USA; n = `r n_dendro_trees_HF` trees from 1998-2003), we fit a logistic growth model[@mcmahon_general_2015] to determine the days of year (DOY) when 25, 50, and 75% annual growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$), peak growing season length ($L_{pgs}= DOY_{75}$-$DOY_{25}$), the maximum daily growth rates ($g_{max}$) and the DOY on which it occurred ($DOY_{g_{max}}$), and total annual increment in diameter at breast height ($\Delta DBH$; Fig. `r schematic_summary`). 
This analysis was performed separately for ring- (69 trees at SCBI and 138 trees at Harvard Forest) and diffuse-porous (54 trees at SCBI and 202 trees at Harvard Forest) species, which differ in growth phenology[@dorangeville_peak_2021]. These stem-growth milestones were compared to canopy foliage phenology (measured at ecosystem level via remote sensing).

>>>>

While responses of leaf phenology and seasonal _NEE_ to warming spring temperatures have been documented[@jeong_phenology_2011; @keenan_net_2014; @buermann_widespread_2018; @ibanez_forecasting_2010; @menzel_european_2006; @menzel_growing_1999;@finzi_carbon_2020], little is known about how the longest-lived component of fixed C in trees, the woody growth, is responding to warming spring temperatures. 
In fact, we are aware of only a few studies that have documented stem-growth phenology of temperate deciduous forests over multiple years[@dorangeville_peak_2021; @etzold_number_2022; @tumajer_limitation_2022]. <!-- I'm not aware of any others, but we do want to make sure we're not missing anything. klesse_integrating_2016 and martin_hydrometeorology_2017 both do this, but for conifers.-->
**The climate sensitivity of woody growth phenology in temperate deciduous trees and its link to annual growth has never been studied *in-situ* (but see Ref.[@zohner_how_2021] for a controlled sapling experiment).** <!-- revise this statement. Etzold et al kind of do this.-->

@@

Using dendrometer band measurements taken throughout multiple growing seasons at the Smithsonian Conservation Biology Institute (SCBI; Virginia, USA; n = `r n_dendro_trees_SCBI` trees from 2011-2020) and Harvard Forest (Massachusetts, USA; n = `r n_dendro_trees_HF` trees from 1998-2003), we fit a logistic growth model[@mcmahon_general_2015] to determine the days of year (DOY) when 25, 50, and 75% annual growth were achieved ($DOY_{25}$, $DOY_{50}$, $DOY_{75}$), peak growing season length ($L_{pgs}= DOY_{75}$-$DOY_{25}$), the maximum daily growth rates ($g_{max}$) and the DOY on which it occurred ($DOY_{g_{max}}$), and total annual increment in diameter at breast height ($\Delta DBH$; Fig. `r schematic_summary`). 
This analysis was performed separately for ring- (69 trees at SCBI and 138 trees at Harvard Forest) and diffuse-porous (54 trees at SCBI and 202 trees at Harvard Forest) species, which differ in growth phenology[@dorangeville_peak_2021; @tumajer_limitation_2022]. These stem-growth milestones were compared to canopy foliage phenology (measured at ecosystem level via remote sensing).
